---
title: "Perception of groove: Simulation and Validation"
author: "Ruta Slivkaite"
date: "2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading libraries
```{r}
library(rjags)
library(coda)
library(jsonlite)

```

## Loading data
```{r, echo=FALSE}

# Set working directory
setwd("C:/Users/Ruta/Documents/Rutos SmuTkes/CogSci/Master/Decision making")

# Load data
MRAnalysis <- read.csv("Data/Data_MRAnalysis_OSF.csv") # all variables for the analysis

Groove_Ratings <- read.csv("Data/Groove_Ratings_Data_OSF.csv") # groove ratings (objective)


# Load Metadata
json_data_Groove_Ratings <- fromJSON("Data/a4ed8-datacite.json") # all variables for the analysis

json_data_MRAnalysis <- fromJSON("Data/cnmv5-datacite.json") # groove ratings


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Data pre-processing
```{r}
mean(MRAnalysis$GOLD_MSI_MusicalSophistication)
sd(MRAnalysis$GOLD_MSI_MusicalSophistication)

mean(MRAnalysis$GOLD_DSI_ParticipatoryDanceExperience)
sd(MRAnalysis$GOLD_DSI_ParticipatoryDanceExperience)

hist(MRAnalysis$GOLD_MSI_MusicalSophistication, freq = FALSE, main = "Histogram and Density Plot")
# Q-Q plot
qqnorm(MRAnalysis$GOLD_MSI_MusicalSophistication)
qqline(MRAnalysis$GOLD_MSI_MusicalSophistication, col = "red")

mean(MRAnalysis$HG_groove)
mean(MRAnalysis$LG_groove)

mean(MRAnalysis$HG_familiar)
sd(MRAnalysis$HG_familiar)

mean(MRAnalysis$LG_familiar)
sd(MRAnalysis$LG_familiar)

mean(MRAnalysis$HG_like)
sd(MRAnalysis$HG_like)

mean(MRAnalysis$LG_like)
sd(MRAnalysis$LG_like)

mean(MRAnalysis$BeatSensitivity)
sd(MRAnalysis$BeatSensitivity)

mean(MRAnalysis$MeterSensitivity)
sd(MRAnalysis$MeterSensitivity)

hist(MRAnalysis$LG_like, freq = FALSE, main = "Histogram and Density Plot")

```



## Modeling with RJAGS
### Parameters recovery

#### MODEL 
#### Step 1: JAGS model specification

```{r}

model {
  for (i in 1:N) { # N = number of songs
    is_high_groove[i] <- equals(grooveLevel[i], 1)
    
    for (j in 1:P) { # P = number of participants
      
      # Sub-model for Familiarity as influenced by MSI
      Familiarity[i, j] ~ dnorm(mu_Familiarity[i, j], tau_fam)
      mu_Familiarity[i, j] <- gamma0 + gamma1 * MSI[i, j]

      # Pleasure Rating Model for High Groove Songs
      lambda_high[i, j] <- exp(beta0_high + 
                               beta1_high * objGrooveRating[i] +
                               beta2_high * MSI[i, j] +
                               beta3_high * Familiarity[i, j] +
                               # Interaction term example: Objective Groove Rating * MSI
                               beta4_high * objGrooveRating[i] * MSI[i, j] +
                               beta5_high * DSI[i, j] +
                               beta6_high * objGrooveRating[i] * DSI[i, j] +
                               participant_effect[j]) # random effects
      rating_high[i, j] ~ dpois(lambda_high[i, j]) 
      truncation_high[i, j] <- step(rating_high[i, j] - 1) # Zero-Truncated Poisson

      # Pleasure Rating Model for Low Groove Songs
      lambda_low[i, j] <- exp(beta0_low + 
                              beta1_low * objGrooveRating[i] +
                              beta2_low * MSI[i, j] +
                              beta3_low * Familiarity[i, j] +
                              beta4_low * DSI[i, j] +
                              participant_effect[j]) # random effects
      rating_low[i, j] ~ dpois(lambda_low[i, j]) 
      truncation_low[i, j] <- step(rating_low[i, j] - 1) # Zero-Truncated Poisson

      # Observations using an indicator variable
      rating[i, j] ~ dpois(is_high_groove[i] * lambda_high[i, j] + (1 - is_high_groove[i]) * lambda_low[i, j])
    }
  }
}





```



#### Step 2: Simulation
##### 2.1 Set-up

```{r}

# Set parameters for simulation
N <- 20  # Number of songs
P <- 124  # Number of participants

################### Trues parameters ###################

# Set true parameter values for Familiarity sub-model
true_gamma0 <- 0.2 # baseline
true_gamma1 <- 0.1  # effect of MSI on Familiarity

# Set true beta coefficients for the high and low groove pleasure models
true_beta0_high <- 0.1 # 1.1 # baseline for high groove
true_beta1_high <- 0.2 # 0.2  # effect of objGrooveRating on high groove pleasure
true_beta2_high <- 0.1 # 0.1  # effect of MSI on high groove pleasure
true_beta3_high <- 0.2 # 0.2   # effect of Familiarity on high groove pleasure
true_beta4_high <- 0.05 #0.05    # interaction MSI and objective rating:
true_beta5_high <- 0.05 # 0.05    # effect of DSI on high groove pleasure


true_beta0_low <- -0.5 #0.6  # baseline for low groove
true_beta1_low <- 0.01 # 0.01   # effect of objGrooveRating on low groove pleasure
true_beta2_low <- -0.6 # 0.54 effect of MSI on low groove pleasure
true_beta3_low <- 0.01 # 0.01  # effect of Familiarity on low groove pleasure
true_beta4_low <- -0.6    # effect of DSI on low groove pleasure
# No direct effect of Familiarity for low groove songs


# Set true values for the precision of the participant random effects
true_tau_participant <- 1.0

# Set true values for the precision of the familiarity model
true_tau_fam <- 1.5  




################### Simulate data ###################
set.seed(57)


# Objective groove rating for each song
#preset_objGrooveRatings <- c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.01, 0.001, 0.002, 0.003, 0.004, 0.005, 0.006, 0.007, 0.008, 0.009, 0.0001)

preset_objGrooveRatings <- c(108.7, 105.9, 105.1, 102.5, 102.4, 101.6, 101.1, 98.7, 97.4, 96.9, 38.7, 38.5, 38.1, 36.5, 35.2, 34.8, 32.6, 32.1, 30.9, 29.3)
min_rating <- min(preset_objGrooveRatings)
max_rating <- max(preset_objGrooveRatings)
scaled_objGrooveRatings <- (preset_objGrooveRatings - min_rating) / (max_rating - min_rating)

N_songs <- 20  # Total number of songs
grooveLevel <- c(rep(1, N_songs/2), rep(0, N_songs/2))
# Shuffle the objGrooveRating in the same order as grooveLevel
objGrooveRating <- scaled_objGrooveRatings[order(match(grooveLevel, sample(c(0, 1), N, replace = TRUE)))]


# Generate other variables
#MSI <- matrix(rnorm(N * P, mean = 73.5, sd = 17.9), nrow = N, ncol = P)  # MSI
#DSI <- matrix(rnorm(N * P, mean = 83.8 , sd = 23.7 ), nrow = N, ncol = P)  # DSI
# MSI
R <- 20 # rows
C <- 124  # columns
# Generate a random number for each column
random_values <- rnorm(C,  mean = 73.5, sd = 17.9)
# Replicate each number N times and create the matrix
MSI <- matrix(rep(random_values, each = R), nrow = R, ncol = C)
# Calculating the mean and standard deviation of the matrix
mean_MSI <- mean(MSI)
sd_MSI <- sd(MSI)
# Standardizing the matrix
MSI <- (MSI - mean_MSI) / sd_MSI



# DSI
R <- 20 # rows
C <- 124  # columns
# Generate a random number for each column
random_values_DSI <- rnorm(C,  mean = 83.8, sd = 23.7)
# Replicate each number N times and create the matrix
DSI <- matrix(rep(random_values_DSI, each = R), nrow = R, ncol = C)
# Calculating the mean and standard deviation of the matrix
mean_DSI <- mean(DSI)
sd_DSI <- sd(DSI)
# Standardizing the matrix
DSI <- (DSI - mean_DSI) / sd_DSI


#MSI <- matrix(runif(N * P, min = 0, max = 10), nrow = N, ncol = P)  # Rescaled MSI
#DSI <- matrix(runif(N * P, min = 0, max = 10), nrow = N, ncol = P)  # Rescaled DSI

# Familiarity
Familiarity <- matrix(sample(1:7, 20 * 124, replace = TRUE), nrow = 20, ncol = 124)

participant_effect <- rnorm(P, 0, sqrt(1 / true_tau_participant))  # Participant effect

# Initialize matrices for pleasure ratings
rating_high <- matrix(nrow = N, ncol = P)
rating_low <- matrix(nrow = N, ncol = P)
# Initialize matrices for lambda values
lambda_high <- matrix(nrow = N, ncol = P)
lambda_low <- matrix(nrow = N, ncol = P)



# Simulate pleasure ratings based on the model
for (i in 1:N) {
  for (j in 1:P) {
    
    # Familiarity model
    mu_Familiarity <- true_gamma0 + true_gamma1 * MSI[i, j]
    Familiarity[i, j] <- rnorm(1, mu_Familiarity, sqrt(1 / true_tau_fam))

    # High groove pleasure rating model
    lambda_high[i, j] <- exp(true_beta0_high + 
                         true_beta1_high * objGrooveRating[i] + 
                         true_beta2_high * MSI[i, j] +  
                         true_beta3_high * Familiarity[i, j] + 
                         true_beta4_high * objGrooveRating[i] * MSI[i, j] +
                         true_beta5_high * DSI[i, j] + 
                         participant_effect[j])
    
    #rating_high[i, j] <- rpois(1, lambda_high)
    #rating_high[i, j] <- ifelse(is.finite(lambda_high[i, j]) && lambda_high[i, j] > 0, rpois(1, lambda_high[i, j]), NA)

    rating_high[i, j] <- min(max(rpois(1, lambda_high[i, j]), 1), 7)

    
    
    # Low groove pleasure rating model
    lambda_low[i, j] <- exp(true_beta0_low + 
                      true_beta1_low * objGrooveRating[i] +
                      true_beta2_low * MSI[i, j] + 
                      true_beta3_low * Familiarity[i, j] +  
                      true_beta4_low * DSI[i, j] + 
                      participant_effect[j])
    
    #rating_low[i, j] <- rpois(1, lambda_low)
    #rating_low[i, j] <- ifelse(is.finite(lambda_low[i, j]) && lambda_low[i, j] > 0, rpois(1, lambda_low[i, j]), NA)
    rating_low[i, j] <- min(max(rpois(1, lambda_low[i, j]), 1), 7)
  }
}

# Combine high and low groove ratings based on groove level
rating <- matrix(nrow = N, ncol = P)
for (i in 1:N) {
  for (j in 1:P) {
    rating[i, j] <- ifelse(grooveLevel[i] == 1, rating_high[i, j], rating_low[i, j])
  }
}


```


##### 2.2 Running simulation
```{r}
set.seed(77)
################### Prepare data for JAGS ###################
data_jags <- list(
  N = N,
  P = P,
  rating = rating,
  objGrooveRating = objGrooveRating,
  MSI = MSI,
  DSI = DSI,
  Familiarity = Familiarity,
  grooveLevel = grooveLevel,
  participant_effect = participant_effect
  
)

# Load the model
model_file = "Models/GROOVE_model_Sim.txt"
jags_model = jags.model(file = model_file, data = data_jags, n.chains = 3, n.adapt = 1000)

# Burn-in period
update(jags_model, 1000)

# Sample from the posterior
samples_sim = coda.samples(jags_model, variable.names = c("gamma0", "gamma1","beta0_high", "beta1_high", "beta2_high", "beta3_high", "beta4_high", "beta5_high", "beta0_low", "beta1_low", "beta2_low", "beta3_low", "beta4_low",  "tau_participant_effect", "tau_fam"), n.iter = 8000)


# Save the MCMC samples to a file
save(samples_sim, file = "mcmc_samples_simulation.RData")

```

##### 2.3 Results
###### a) Estimates

```{r}


# View the results
print(summary(samples_sim))
exp_results_sim <- exp(summary(samples_sim)$statistics[, "Mean"])
print(exp_results_sim)





```

```{r}
exp(0.11)
```



```{r}



# Create trace plots
traceplot(samples_sim)




```

```{r, save trace plots}

library(ggplot2)
library(coda)
library(dplyr)
library(tidyr)

# Extract data from the mcmc.list object
chain_list <- lapply(1:length(samples_sim), function(x) {
  as.data.frame(samples_sim[[x]])
})

# Add chain and iteration information
for (i in 1:length(chain_list)) {
  chain_list[[i]]$Chain <- i
  chain_list[[i]]$Iteration <- 1:nrow(chain_list[[i]])
}

# Combine all chains into one data frame
df_samples <- bind_rows(chain_list)

# Convert to long format suitable for ggplot
long_samples <- gather(df_samples, key = "Parameter", value = "Value", -Chain, -Iteration)

# Create the plot
p <- ggplot(long_samples, aes(x = Iteration, y = Value, color = as.factor(Chain))) +
    geom_line() +
    facet_wrap(~ Parameter, scales = "free")

# Save the plot
ggsave("traceplot_simulation.png", plot = p, width = 10, height = 8)



```





###### b) Density plots
```{r}


library(ggplot2)


# Combine all chains into a single matrix
sample_matrix = do.call(rbind, samples_sim)

# Convert the matrix to a data frame
sample_df = as.data.frame(sample_matrix)


########### Density plots for beta´s in High Groove category ###########

# Density plot for beta0_high 
p1 <- ggplot(sample_df, aes(x = beta0_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.1, color = "red", linetype = "dashed") +
    ggtitle("Baseline High Groove PLUMM Rating") +
    xlab("beta0_high") +
    ylab("Density") +
    theme_classic()
p1


# Density plot for beta1_high
p2 <- ggplot(sample_df, aes(x = beta1_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.2, color = "red", linetype = "dashed") +
    ggtitle("Effect of Objective Groove Rating on High Groove PLUMM Rating") +
    xlab("beta1_high") +
    ylab("Density")+
    theme_classic()



# Density plot for beta2_high
p3 <- ggplot(sample_df, aes(x = beta2_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.1, color = "red", linetype = "dashed") +
    ggtitle("Effect of MSI on High Groove PLUMM Rating") +
    xlab("beta2_high") +
    ylab("Density")+
    theme_classic()



# Density plot for beta3_high
p4 <- ggplot(sample_df, aes(x = beta3_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.2, color = "red", linetype = "dashed") +
    ggtitle("Effect of Familiarity on High Groove PLUMM Rating") +
    xlab("beta3_high") +
    ylab("Density")+
    theme_classic()


# Density plot for beta4_high
p5 <- ggplot(sample_df, aes(x = beta4_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
    ggtitle("Interaction between Objective Groove Rating and MSI") +
    xlab("beta4_high") +
    ylab("Density")+
    theme_classic()



# Density plot for beta5_high
p6 <- ggplot(sample_df, aes(x = beta5_high)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.05, color = "red", linetype = "dashed") +
    ggtitle("Effect of DSI on High Groove PLUMM Rating") +
    xlab("beta5_high") +
    ylab("Density")+
    theme_classic()






########### Density plots for beta´s in Low Groove category ###########

# Density plot for beta0_low
p8 <- ggplot(sample_df, aes(x = beta0_low)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = -0.5, color = "red", linetype = "dashed") +
    ggtitle("Baseline Low Groove PLUMM Rating") +
    xlab("beta0_low") +
    ylab("Density") +
    theme_classic()
p8



# Density plot for beta1_low
p9 <- ggplot(sample_df, aes(x = beta1_low)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.01, color = "red", linetype = "dashed") +
    ggtitle("Effect of Objective Groove Rating on Low Groove PLUMM Rating") +
    xlab("beta1_low") +
    ylab("Density") +
    theme_classic()



# Density plot for beta2_low
p10 <- ggplot(sample_df, aes(x = beta2_low)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = -0.6, color = "red", linetype = "dashed") +
    ggtitle("Effect of MSI on Low Groove PLUMM Rating") +
    xlab("beta2_low") +
    ylab("Density")+
    theme_classic()



# Density plot for beta3_low
p11 <- ggplot(sample_df, aes(x = beta3_low)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.01, color = "red", linetype = "dashed") +
    ggtitle("Effect of Familiarity on Low Groove PLUMM Rating") +
    xlab("beta3_low") +
    ylab("Density")+
    theme_classic()


# Density plot for beta4_low
p12 <- ggplot(sample_df, aes(x = beta4_low)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = -0.6, color = "red", linetype = "dashed") +
    ggtitle("Effect of DSI on Low Groove PLUMM Rating") +
    xlab("beta4_low") +
    ylab("Density")+
    theme_classic()


########## Plot gamma as well #########

# Density plot for gamma0
p13 <- ggplot(sample_df, aes(x = gamma0)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.2, color = "red", linetype = "dashed") +
    ggtitle("Familiarity baseline") +
    xlab("gamma0") +
    ylab("Density")+
    theme_classic()
p13

# Density plot for gamma1
p14 <- ggplot(sample_df, aes(x = gamma1)) +
    geom_density(fill = "blue", alpha = 0.5) +
    geom_vline(xintercept = 0.1, color = "red", linetype = "dashed") +
    ggtitle("Effect of MSI on Familiarity") +
    xlab("gamma1") +
    ylab("Density")+
    theme_classic()
p14

# Set true parameter values for Familiarity sub-model
true_gamma0 <- 0.2 # baseline
true_gamma1 <- 0.1  # effect of MSI on Familiarity



```

###### bb) Combined density plots
```{r}

library(gridExtra)

# Combined density plots

## For baseline levels
#p1 and p8
p1_p8 <- grid.arrange(p1, p8, ncol = 2)
# Save the combined plot
ggsave("baseline_density_plots.png", plot = p1_p8, width = 10, height = 5)


## For MSI and DSI
#p3, p6 and p10, p12
p3_p6_p10_p12 <- grid.arrange(p3, p6, p10, p12, ncol = 2)
# Save the combined plot
ggsave("MSI_DSI_density_plots.png", plot = p3_p6_p10_p12, width = 10, height = 5)


## For Interaction terms in High groove
p5 <- grid.arrange(p5, ncol = 2)
# Save the combined plot
ggsave("HighGroove_interactions_density_plots.png", plot = p5, width = 10, height = 5)


## For familiarity
p4_p11 <- grid.arrange(p4, p11, ncol = 2)
# Save the combined plot
ggsave("Familiarity_density_plots.png", plot = p4_p11, width = 10, height = 5)

## For objective groove rating
p2_p9 <- grid.arrange(p2, p9, ncol = 2)
# Save the combined plot
ggsave("ObjGrooveRating_density_plots.png", plot = p2_p9, width = 10, height = 5)

## For familiarity
p13_p14 <- grid.arrange(p13, p14, ncol = 2)
# Save the combined plot
ggsave("Familiarity_density_plots.png", plot = p13_p14, width = 10, height = 5)

combined <- grid.arrange(p1_p8, p3_p6_p10_p12,p5, p4_p11, p2_p9,  p13, p14, ncol = 2)
# Save the combined plot
ggsave("All_density_plots.png", plot = combined, width = 15, height = 20)


```


###### c) Scatter plots

```{r}


## Plot estimated vs true value

true_values <- c( beta0_high = 0.1, beta0_low = -0.5,  beta1_high = 0.2, beta1_low = 0.01, beta2_high = 0.1, beta2_low = -0.6, beta3_high = 0.2, beta3_low = 0.01, beta4_high = 0.05, beta4_low = -0.6, beta5_high = 0.05, gamma0 = 0.2, gamma1 = 0.1,  tau_fam = 1.5, tau_participant = 1.0)

# Extracting the mean of the estimated parameters
estimated_means <- as.data.frame(summary(samples_sim)$statistics)[, "Mean"]
# allgnign the names
names(estimated_means) <- names(true_values)


# Plotting true vs estimated values
plot_data <- data.frame(True = true_values, Estimated = estimated_means)
p15 <- ggplot(plot_data, aes(x = True, y = Estimated, label = c( "beta0_high", "beta0_low",  "beta1_high", "beta1_low","beta2_high", "beta2_low", "beta3_high","beta3_low", "beta4_high", "beta4_low", "beta5_high", "gamma0", "gamma1", "tau_participant_effect", "tau_fam"))) +
    geom_point() +
    geom_text(nudge_y = 0.2, size = 2) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
    ggtitle("True vs Estimated Parameter Values") +
    xlab("True Values") +
    ylab("Estimated Means") +
    theme_classic()

# save the plot
ggsave("true vs estimated values.png", plot = p15, width = 10, height = 5)


## plot random effects
participant_effects_samples <- as.matrix(samples_sim)[, grep("tau_participant_effect", colnames(as.matrix(samples_sim)))]
dim(participant_effects_samples)
participant_effects_samples
participant_effects_samples<- as.matrix(participant_effects_samples)
dim(participant_effects_samples)

# Calculate the mean of the posterior distribution for each participant's effect
estimated_participant_effects <- apply(participant_effects_samples, 2, mean)
estimated_participant_effects

plot_data <- data.frame( 
                        ParticipantEffect = estimated_participant_effects)

plot_data$MSIScore <- msi_scores

msi_scores <- MSI[1, ]
msi_scores_df <- data.frame(msi_scores)
msi_scores_df

ggplot(plot_data, aes(x = msi_scores_df$msi_scores, y = ParticipantEffect)) +
    geom_point() +
    ggtitle("Random Effects vs. MSI Scores") +
    xlab("MSI Score") +
    ylab("Estimated Random Effect")

participant_effects_samples

```


### Model validation

#### Preparing the dataset

```{r}

# Loading full dataset
MRAnalysis <- read.csv("Data/Data_MRAnalysis_OSF.csv") # all variables for the analysis

Groove_Ratings <- read.csv("Data/Groove_Ratings_Data_OSF.csv") # groove ratings (objective)

########### Preparing MSI and DSI ###########

# Extracting necessary columns
# GOLD_MSI_MusicalSophistication, GOLD_DSI_DanceTraining
MSI_val <- MRAnalysis[c("Participants", "GOLD_MSI_MusicalSophistication")]
DSI_val <- MRAnalysis[c("Participants", "GOLD_DSI_DanceTraining")]
# arrange by Participant in a descending order and get MSI and DSI
MSI_sorted <- MSI_val %>% arrange(Participants)
DSI_sorted <- DSI_val %>% arrange(Participants)


# Convert from long to wide format
# MSI
wide_MSI <- MSI_sorted %>% 
  pivot_wider(names_from = Participants, values_from = c("GOLD_MSI_MusicalSophistication"))
MSI_val <- wide_MSI[rep(1, 20), ]
# DSI
wide_DSI <- DSI_sorted %>% 
  pivot_wider(names_from = Participants, values_from = c("GOLD_DSI_DanceTraining"))
DSI_val <- wide_DSI[rep(1, 20), ]

# Transform it into a matrix
MSI_val_matrix <- as.matrix(MSI_val)
DSI_val_matrix <- as.matrix(DSI_val)

# Normalize MSI and DSI
##### MSI
# Calculating the mean and standard deviation of the matrix
mean_MSI_val <- mean(MSI_val_matrix)
sd_MSI_val <- sd(MSI_val_matrix)
# Standardizing the matrix
MSI_val_matrix <- (MSI_val_matrix - mean_MSI_val) / sd_MSI_val

###### DSI
# Calculating the mean and standard deviation of the matrix
mean_DSI_val <- mean(DSI_val_matrix)
sd_DSI_val <- sd(DSI_val_matrix)
# Standardizing the matrix
DSI_val_matrix <- (DSI_val_matrix - mean_DSI_val) / sd_DSI_val





########### Preparing Groove, Familiarity, Pleasure ratings ###########

# Loading participant ratings for each song
Participant_Song_Ratings <- read.csv("Data/Participant_Song_Ratings.csv", header = TRUE, sep = ";")

# Separate like, groove, familiarity. Reverse it so each song is in the rows (20 rows) and every participant rating is in the columns

# Number of songs
num_songs <- 20

# Create indices for each rating type
groove_indices <- seq(1, num_songs * 3, by = 3)
like_indices <- seq(2, num_songs * 3, by = 3)
familiar_indices <- seq(3, num_songs * 3, by = 3)

# Extract columns for each rating type
groove_indices <- grep("\\.groove$", names(Participant_Song_Ratings))
like_indices <- grep("\\.like$", names(Participant_Song_Ratings))
familiar_indices <- grep("\\.familiar$", names(Participant_Song_Ratings))

participant_id_col <- 1
# Extract columns for each rating type
groove_df<- Participant_Song_Ratings[, c(participant_id_col, groove_indices)]
like_df <- Participant_Song_Ratings[, c(participant_id_col, like_indices)]
familiar_df <- Participant_Song_Ratings[, c(participant_id_col, familiar_indices)]


####### Familiarity #######
## Create a matrix where each row is a song and each column is a value how familiar participant is with that song
# Transpose the data
transposed_familiar <- t(familiar_df)
# Convert back to data frame 
familiar_df <- as.data.frame(transposed_familiar)
# Set the names of the columns to be the values in the first row
colnames(familiar_df) <- as.character(familiar_df[1, ])
# Remove the first row 
familiar_df <- familiar_df[-1, ]
# Transform it into a matrix
familiar_matrix <- as.matrix(familiar_df)


####### Pleasure  #######
# Transpose the data
transposed_like <- t(like_df)
# Convert back to data frame 
like_df <- as.data.frame(transposed_like)
# Set the names of the columns to be the values in the first row
colnames(like_df) <- as.character(like_df[1, ])
# Remove the first row 
like_df <- like_df[-1, ]
# Transform it into a matrix
like_matrix <- as.matrix(like_df)


####### Groove  #######
# Transpose the data
transposed_groove <- t(groove_df)
# Convert back to data frame 
groove_df <- as.data.frame(transposed_groove)
# Set the names of the columns to be the values in the first row
colnames(groove_df) <- as.character(groove_df[1, ])
# Remove the first row 
groove_df <- groove_df[-1, ]
# Transform it into a matrix
groove_matrix <- as.matrix(groove_df)


####### Objective groove ratings #######

preset_objGrooveRatings <- c(108.7, 105.9, 105.1, 102.5, 102.4, 101.6, 101.1, 98.7, 97.4, 96.9, 38.7, 38.5, 38.1, 36.5, 35.2, 34.8, 32.6, 32.1, 30.9, 29.3)
min_rating <- min(preset_objGrooveRatings)
max_rating <- max(preset_objGrooveRatings)
scaled_objGrooveRatings <- (preset_objGrooveRatings - min_rating) / (max_rating - min_rating)

N_songs <- 20  # Total number of songs
grooveLevel <- c(rep(1, N_songs/2), rep(0, N_songs/2))
# Organize the objGrooveRating in the same order as grooveLevel
objGrooveRating <- scaled_objGrooveRatings[order(match(grooveLevel, sample(c(0, 1), N, replace = TRUE)))]

```


#### Running the model
```{r}

set.seed(88)
N <- 20  # Number of songs
P <- 124  # Number of participants

################### Prepare data for JAGS ###################
data_jags_val <- list(
  N = N,
  P = P,
  rating = like_matrix, # using like variable
  objGrooveRating = objGrooveRating,
  MSI = MSI_val_matrix,
  DSI = DSI_val_matrix,
  Familiarity = familiar_matrix,
  grooveLevel = grooveLevel,
  participant_effect = participant_effect
  
)

# Load the model
model_file = "Models/GROOVE_model_Sim.txt"
jags_model_val = jags.model(file = model_file, data = data_jags_val, n.chains = 3, n.adapt = 1000)

# Burn-in period
update(jags_model_val, 1000)

# Sample from the posterior
samples_val = coda.samples(jags_model_val, variable.names = c("gamma0", "gamma1","beta0_high", "beta1_high", "beta2_high", "beta3_high", "beta4_high", "beta5_high", "beta0_low", "beta1_low", "beta2_low", "beta3_low", "beta4_low", "tau_participant_effect", "tau_fam"), n.iter = 8000)


# Save the MCMC samples to a file
save(samples_val, file = "mcmc_samples_validation.RData")





```



```{r}


samples <- coda.samples(jags_model_val, variable.names = c("participant_effect"), n.iter = 1000)
# Extracting the participant effects
estimated_effects <- as.matrix(samples_val)[, "participant_effect"]
```


```{r}

# Create trace plots
traceplot(samples_val)



```




```{r}



# View the results
print(summary(samples_val))
# Create trace plots
traceplot(samples_val)



```

